---
title: "TAIC Legacy report example report - Breadth"
author: "James Thompson"
date: "`r Sys.Date()`"
abstract: "This is an example report for the TAIC Legacy project. It takes the summary.csv file and does normal EDA on it. This should not be considered an extensive analysis of the data created by the engine. Simply a demonstration of things you might do with it."
output: pdf_document
geometry: margin=1cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6, fig.height = 4, fig.align = "center")
options(knitr.table.format = "latex")

library(tidyverse)
library(factoextra)

# Load data
data <- read_csv("output/summary.csv")

# Create commonly used derivative data frames
data_long <- data |>
    pivot_longer(cols = !ReportID, names_to = "theme", values_to = "weighting")

data_themes_only <- data |>
    dplyr::select(-ReportID)
```

# Introduction

The engine uses AI to read through all the marine accident investigations reports and extract the common themes and provide weightings to each of the themes.

This means that the two main outputs of the engine are overall themes and weightings for each report.

To help visualize the output from the engine an online [viewer app](https://taic-viewer-72e8675c1c03.herokuapp.com/).

# Data

The data we are dealing with are about 50 reports retrieved from the public taic [website](https://www.taic.org.nz/inquiries?order=field_publication_date_value&sort=desc&keyword=&date_filter%5Bmin%5D%5Bdate%5D=&date_filter%5Bmax%5D%5Bdate%5D=&publication_date%5Bmin%5D%5Bdate%5D=&publication_date%5Bmax%5D%5Bdate%5D=&status%5B0%5D=12). A few reports have been discarded because they are either too long or are not in a format that the engine  can read. These are either prelimenary reports or reports that were made in conjuction in another country agency.

# Themes

The 5 themes indentified by the engine are:

```{r}
themes_yaml <- yaml::read_yaml("output/themes.yaml")
themes <- themes_yaml$themes

```

# Weightings


# Conclusion

Overall we can see that there is some interesting analysis that can be done on this data. However, there is also alot of work that needs to be done to make it more useful. For example, the themes are not very well defined and there is alot of variation in the data.

Adding more variables to the data will help with deeper analysis. Currently theme weightings are void of any context. Adding in the actual context of the accident will help with this. For example whether the accident was fatal. This will allow us to do more interesting analysis like "What themes are most common in fatal accidents?".

